---
title: "Homework 2"
author: Gauri Bhatkhande 
output: github_document
editor_options: 
  chunk_output_type: console
---


```{r setup}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1 

Read the Mr. Trashwheel dataset 
```{r}
trashwheel_df = 
  read_xlsx("./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
drop_na(dumpster)%>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017
```{r}
precip_2018 = 
  read_excel(
    "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)
  
precip_2017 = 
  read_excel(
    "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)
  

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Balitimore, Maryland. The dataset contains information on year, month, trash collected, specific kinds of trash. There are total `r nrow(trashwheel_df)` rows. Additional data sheets include month precipitation data. 

## Problem 2 

Reading the NYC Transit data from the csv and cleaning it : 
```{r}
nyc_transit_df =
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(division:route11, entry, vending, entrance_type, ada)

# Converting the entry variable from character to logical  (TRUE, FALSE)
nyc_transit_df$entry =
  ifelse(nyc_transit_df$entry=="YES",TRUE,FALSE )
```

### Short note on the dataset

The given dataset contains the variables: *`r variable.names(nyc_transit_df)`*. So far, I imported the data and read the data from the csv file usinf read_csv function. After that, using the pipe operator I went on to clean the dataset which converted the variable names to lower snake case. For this purpose, I used a function (clean_names) without loading the entire package (janitor). After this, I selected the required column names specified in the homework question using the select function and then selecting the variables in the specified order. In the next step, I converted the entry variable from character to logical  (TRUE, FALSE) using the ifelse function wherein I change the entry variable from nyc_transit_df dataset and assign TRUE for the values that earlier said "yes" and else FALSE. 

The dataset now has `r nrow(nyc_transit_df)` rows and `r ncol(nyc_transit_df)` columns. 

These data are not tidy as the routes are all written as separate variables when they should have been a part of the same varaible or column. 

**Finding number of distinct stations**

*The number of distinct stations are `r count(distinct(nyc_transit_df,paste(nyc_transit_df$line, nyc_transit_df$station_name)))`.*

**Finding how many stations are ADA compliant**

```{r}

ada_compliant_stations = 
  filter(nyc_transit_df, ada==TRUE) %>%
  distinct(paste(line, station_name)) %>%
   count()

ada_compliant_stations
```
* `r filter(nyc_transit_df, ada==TRUE) %>% distinct(paste(line, station_name)) %>% count()` stations are ADA compliant.*

**Finding proportion of station entrances / exits without vending allow entrance**

```{r}
total_entry =
  filter(nyc_transit_df, vending=='NO') %>%
  count()

tot_entry_allowed = 
  filter(nyc_transit_df, vending=='NO') %>%
  filter( entry==TRUE) %>%
  count()


proportion = tot_entry_allowed/total_entry

proportion

```
*The proportion of station entrances / exits without vending that allow entrance is 0.37705 or 37.705%*

**Reformat data so that route number and route name are distinct variables** 

```{r}
nyc_transit_df = mutate(nyc_transit_df,
       route8 = as.character(route8)) %>%
mutate(route9 = as.character(route9)) %>%
mutate(route10 = as.character(route10)) %>%
mutate(route11 = as.character(route11))

nyc_transit_df =
 pivot_longer(
    nyc_transit_df, 
    route1:route11,
    names_to = "route_number", 
    values_to = "route_names")
```


**Finding many distinct stations serve the A train**

```{r}
stations_serve_A = 
  filter(nyc_transit_df, route_names=="A") %>%
  distinct(paste(line, station_name)) %>%
   count()
```
*60 distinct stations serve the A train.*

**Finding how many of the stations that serve the A train are ADA compliant?**

```{r}
stations_serve_A = 
  filter(nyc_transit_df, route_names=="A") %>%
  filter(ada == "TRUE") %>%
  distinct(paste(line, station_name)) %>%
   count()
```

*17 stations that serve the A train are ADA compliant.* 


## Problem 3

**Read data in csv, clean and separate variable mon into integer variables**


```{r}
pols_df =
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-")%>%
  mutate(year = as.numeric(year))%>%
  mutate(month = as.numeric(month))%>%
  mutate(day = as.numeric(day))
  
```

I converted the year and month and day to numeric type so as to facilitate the left join in the following chunk: 
**Replacing month number with month names, creating a president variable taking values gop and dem, and removed prez_dem and prez_gop and the day variable**
month_df had already been created for the Problem 1. I used the same dataframe for the purpose of the following left join; 

```{r}
pols_df=left_join(pols_df,month_df,by="month")%>%
  mutate(month = month_name)%>%
  mutate(president = ifelse(prez_gop==1,"gop","dem"))%>%
  within(rm(prez_gop,prez_dem,day,month_name))

pols_df
```

**Reading and cleaning the snp data and arranging the year and month in ascending order**

```{r}
snp_df =
  read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), sep = "/")%>%
  mutate(year = as.numeric(year))%>%
  mutate(month = as.numeric(month))%>%
  mutate(day = as.numeric(day))%>%
  arrange(year, month)
  
```

**Replacing month number with month names, and organising using relocate function to have year and month as the leading variables** 

```{r}
snp_df=
  left_join(snp_df,month_df,by="month")%>%
  mutate(month = month_name)%>%
  within(rm(day,month_name))%>%
  relocate(year, month)

snp_df
```

**Tidying the unemployment data**

```{r}
unem_df = 
  read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(jan:dec,names_to ="month",values_to="values")
```

**Changing the month names so as to match them with the format in snp and pols data**

```{r}

month_short_df =
  tibble(
    month = c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec"),
    month_name = month.name
  )

unem_df = left_join(unem_df,month_short_df,by = "month") %>%
  mutate(month = month_name)%>%
  within(rm(month_name))
  
unem_df
```

**Merging the datasets using left join**

```{r}
pols_df = left_join(pols_df, snp_df,by=c("year","month"))
pols_df = left_join(pols_df, unem_df,by=c("year","month"))
    
pols_df
```

## Description of the dataset
The dataset pols_df contained information about the year and the party of the president pertaining to that year i.e. democratic or gop. The dataset also contains information about the number of senators, governors and representatives that belonged to one of the two parties pertaining to a certain year since 1947. 

The dataset snp_df contains the variables `r variable.names(snp_df)`. The dataset has the closing values pertaining to every year and every month. 

THe dataset unem_df contains unemployment information for the unemployment percentage for the years and months. It contains the variables `r variable.names(unem_df)`. 

The resulting dataset formed after inner join of pols_df, snp_df and unem_df contains `r variable.names(pols_df)`. The dimensions of the dataset are: 822 x 11. The range of years covered is `r min(pols_df$year)` to `r max(pols_df$year)`. The key variables are year, president (which tells us whether it was a democrat or gop), close, value

